Q1: Are lake and terrestrial primary productivity coherent?

### 2019-07-03 AnalysisCode/ms1_analyses_randforest.Rmd
analysis updates

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)
library(viridis)
library(fields)

@@

hist(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts1<0.2]/pi, main="Accumulated NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvip.ts1<0.2]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

@@

hist(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts2<0.2]/pi, main="Accumulated NDVI, long timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvicoh.ts2>0.6]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

@@

@@

@@

@@

predictors<-analysislakes$lakeinfo
predictors<-left_join(predictors, depth, by="lagoslakeid")
predictors<-left_join(predictors, pct.ag.avg, by="hu12_zoneid")
predictors<-left_join(predictors, avg.chla, by="lagoslakeid")
predictors<-left_join(predictors, tsi.chl, by="lagoslakeid")
predictors<-left_join(predictors, states, by="state_zoneid")
predictors<-left_join(predictors, cv.accndvi, by="lagoslakeid")
#predictors<-left_join(predictors, huc_codes, by="hu4_zoneid")

@@

str(predictors)
hist(predictors$tslength)
hist(predictors$lake_area_ha)
hist(log10(predictors$lake_area_ha))
hist(predictors$lake_perim_meters)
hist(log10(predictors$lake_perim_meters))
table(predictors$nhd_fcode)
table(predictors$hu12_zoneid)
hist(predictors$elevation_m)
hist(predictors$maxdepth)
hist(log10(predictors$maxdepth))
hist(predictors$pct.ag)
hist(log10(predictors$pct.ag))
hist(predictors$chla)
hist(log10(predictors$chla))
hist(predictors$tsi)
table(predictors$tsi.cat)
hist(predictors$cv.accndvi)

predictors$log10_lake_area_ha<-log10(predictors$lake_area_ha)
predictors$log10_lake_perim_meters<-log10(predictors$lake_perim_meters)
predictors$log10_maxdepth<-log10(predictors$maxdepth)
predictors$log10_pct.ag<-log10(predictors$pct.ag+1)
predictors$log10_chla<-log10(predictors$chla)

@@

```{r modelling, echo=TRUE, cache=FALSE}

library(partykit)

cforest.st<-cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, data=modvars.accndvi)
predcoh.st<-predict.cforest(cforest.st, newdata=modvars.accndvi)
hist(predcoh.st)
hist(modvars.accndvi$accndvicoh.ts1)
plot(predcoh.st, modvars.accndvi$accndvicoh.ts1, xlab="predicted", ylab="empirical", main="Coherence, short ts",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.st,modvars.accndvi$accndvicoh.ts1)
varimp.coh.st<-varimp.cforest(cforest.st)
print(varimp.coh.st)

cforest.lt<-cforest(accndvicoh.ts2 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, data=modvars.accndvi)
predcoh.lt<-predict.cforest(cforest.lt, newdata=modvars.accndvi)
hist(predcoh.lt)
hist(modvars.accndvi$accndvicoh.ts2)
plot(predcoh.lt, modvars.accndvi$accndvicoh.ts2, xlab="predicted", ylab="empirical", main="Coherence, long st",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.lt,modvars.accndvi$accndvicoh.ts2)
varimp.coh.lt<-varimp.cforest(cforest.lt)
print(varimp.coh.lt)

cforest.phi.st<-cforest(cos(accndviphi.ts1) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.phist)
predphi.st<-predict.cforest(cforest.phi.st, newdata=modvars.accndvi.phist)
hist(predphi.st)
hist(cos(modvars.accndvi.phist$accndviphi.ts1))
plot(predphi.st, cos(modvars.accndvi.phist$accndviphi.ts1), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.st,cos(modvars.accndvi.phist$accndviphi.ts1))
varimp.phi.st<-varimp.cforest(cforest.phi.st)
print(varimp.phi.st)

cforest.phi.lt<-cforest(cos(accndviphi.ts2) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi + hu4_zoneid + cv.accndvi, 
                        data=modvars.accndvi.philt)
predphi.lt<-predict.cforest(cforest.phi.lt, newdata=modvars.accndvi.philt)
hist(predphi.lt)
hist(cos(modvars.accndvi.philt$accndviphi.ts2))
plot(predphi.lt, cos(modvars.accndvi.philt$accndviphi.ts2), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.lt,cos(modvars.accndvi.philt$accndviphi.ts2))
varimp.phi.lt<-varimp.cforest(cforest.phi.lt)
print(varimp.phi.lt)

@@

>>>>

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)
library(viridis)
library(fields)
#library(partykit)
library(pdp)
library(party)
library(spatstat)

@@

# hist(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts1<0.2]/pi, main="Accumulated NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
rose(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts1<0.3], unit="radian",
     breaks=seq(0,2*pi,length.out=16))
#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvip.ts1<0.2]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

@@

# hist(coh.chlaXaccndvi$accndviphi.ts2[coh.chlaXaccndvi$accndvip.ts2<0.2]/pi, main="Accumulated NDVI, long timescales", xlab="Phase (pi)", ylab="Frequency", breaks=phicls)
rose(coh.chlaXaccndvi$accndviphi.ts2[coh.chlaXaccndvi$accndvip.ts2<0.3], unit="radian",
     breaks=seq(0,2*pi,length.out=16))

#hist(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvicoh.ts2>0.6]/pi, main="Maximum NDVI, short timescales", xlab="Phase (pi)", ylab="Frequency", breaks=5)

@@

#Wetlands
pct.wetlands<-lagosne_select(table="hu12.lulc", vars=c(c("hu12_zoneid","hu12_nlcd2001_pct_90","hu12_nlcd2006_pct_90",
                                                         "hu12_nlcd2011_pct_90","hu12_nlcd2001_pct_95","hu12_nlcd2006_pct_95",
                                                         "hu12_nlcd2011_pct_95")))
pct.wetlands<-pct.wetlands[pct.wetlands$hu12_zoneid %in% analysislakes$lakeinfo$hu12_zoneid,]
pct.wetlands$sum2001<-rowSums(pct.wetlands[,c(2,5)])
pct.wetlands$sum2006<-rowSums(pct.wetlands[,c(3,6)])
pct.wetlands$sum2011<-rowSums(pct.wetlands[,c(4,7)])
pct.wetlands.avg<-data.frame(hu12_zoneid=pct.wetlands$hu12_zoneid, pct.wetlands=rowMeans(pct.wetlands[,8:10]))

@@

#growing season DOC
doc<-lagosne_select(table="epi_nutr", vars=c("lagoslakeid","samplemonth","doc"))
doc<-doc[doc$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
gs.doc<-doc[doc$samplemonth %in% 5:9,]
avg.doc<-aggregate(doc ~ lagoslakeid, data=gs.doc, FUN=mean, na.rm=T)

@@

#mean precipitation
prcp.normal<-raster("~/Box Sync/NSF EAGER Synchrony/Data/PRISM Data/PRISM_ppt_30yr_normal_800mM2_annual_bil.bil")
lakepts<-SpatialPoints(coords=cbind(analysislakes$lakeinfo$nhd_long,analysislakes$lakeinfo$nhd_lat))
lake.prcp<-data.frame(lagoslakeid=analysislakes$lakeinfo$lagoslakeid, prcp.normal=raster::extract(prcp.normal,lakepts))

@@

predictors<-analysislakes$lakeinfo
predictors<-left_join(predictors, depth, by="lagoslakeid")
predictors<-left_join(predictors, pct.ag.avg, by="hu12_zoneid")
predictors<-left_join(predictors, pct.wetlands.avg, by="hu12_zoneid")
predictors<-left_join(predictors, avg.chla, by="lagoslakeid")
predictors<-left_join(predictors, tsi.chl, by="lagoslakeid")
predictors<-left_join(predictors, states, by="state_zoneid")
predictors<-left_join(predictors, cv.accndvi, by="lagoslakeid")
predictors<-left_join(predictors, avg.doc, by="lagoslakeid")
predictors<-left_join(predictors, lake.prcp, by="lagoslakeid")
#predictors<-left_join(predictors, huc_codes, by="hu4_zoneid")

@@

str(predictors)
hist(predictors$tslength)
hist(predictors$lake_area_ha)
hist(log10(predictors$lake_area_ha))
hist(predictors$lake_perim_meters)
hist(log10(predictors$lake_perim_meters))
table(predictors$nhd_fcode)
table(predictors$hu12_zoneid)
hist(predictors$elevation_m)
hist(predictors$maxdepth)
hist(log10(predictors$maxdepth))
hist(predictors$pct.ag)
hist(log10(predictors$pct.ag))
hist(predictors$chla)
hist(log10(predictors$chla))
hist(predictors$tsi)
table(predictors$tsi.cat)
hist(predictors$cv.accndvi)
hist(predictors$pct.wetlands)
hist(predictors$doc)
hist(predictors$prcp.normal)

# predictors$log10_lake_area_ha<-log10(predictors$lake_area_ha) #not necessary to transform with random forests
# predictors$log10_lake_perim_meters<-log10(predictors$lake_perim_meters)
# predictors$log10_maxdepth<-log10(predictors$maxdepth)
# predictors$log10_pct.ag<-log10(predictors$pct.ag+1)
# predictors$log10_chla<-log10(predictors$chla)

@@

```{r modellingCoherence, echo=TRUE, cache=FALSE}

# cforest.st<-partykit::cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + cv.accndvi, 
#                     data=modvars.accndvi, ntree=20000)
cforest.st<-party::cforest(accndvicoh.ts1 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + 
                             cv.accndvi + pct.wetlands + doc + prcp.normal,
                    data=modvars.accndvi, controls=cforest_control(ntree=80000))
predcoh.st<-predict(cforest.st, newdata=modvars.accndvi,type="response")
hist(predcoh.st)
hist(modvars.accndvi$accndvicoh.ts1)
plot(predcoh.st, modvars.accndvi$accndvicoh.ts1, xlab="predicted", ylab="empirical", main="Coherence, short ts",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.st,modvars.accndvi$accndvicoh.ts1)
varimp.coh.st<-varimp(cforest.st)
print(varimp.coh.st[order(varimp.coh.st, decreasing=T)])

partial(cforest.st, pred.var="prcp.normal", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="lake_perim_meters", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="pct.ag", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="lake_area_ha", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="maxdepth", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="chla", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="doc", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="pct.wetlands", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="cv.accndvi", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="tsi.cat", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.st, pred.var="hu4_zoneid", train=modvars.accndvi, type="regression", plot=T)

cforest.lt<-party::cforest(accndvicoh.ts2 ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + 
                             cv.accndvi+ pct.wetlands + doc + prcp.normal, 
                    data=modvars.accndvi, controls=cforest_control(ntree=80000))
predcoh.lt<-predict(cforest.lt, newdata=modvars.accndvi)
# hist(predcoh.lt)
# hist(modvars.accndvi$accndvicoh.ts2)
plot(predcoh.lt, modvars.accndvi$accndvicoh.ts2, xlab="predicted", ylab="empirical", main="Coherence, long st",
     xlim=c(0,1), ylim=c(0,1))
abline(a=0,b=1)
cor.test(predcoh.lt,modvars.accndvi$accndvicoh.ts2)
varimp.coh.lt<-varimp(cforest.lt)
print(varimp.coh.lt[order(varimp.coh.lt, decreasing=T)])

partial(cforest.lt, pred.var="pct.wetlands", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="cv.accndvi", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="hu4_zoneid", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="doc", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="lake_area_ha", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="lake_perim_meters", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="prcp.normal", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="chla", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="tsi.cat", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="pct.ag", train=modvars.accndvi, type="regression", plot=T)
partial(cforest.lt, pred.var="maxdepth", train=modvars.accndvi, type="regression", plot=T)

@@

```{r modellingPhase, echo=TRUE, cache=FALSE}

cforest.phi.st<-cforest(cos(accndviphi.ts1) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + 
                          cv.accndvi+ pct.wetlands + doc + prcp.normal, 
                        data=modvars.accndvi.phist,controls=cforest_control(ntree=40000))
predphi.st<-predict(cforest.phi.st, newdata=modvars.accndvi.phist)
# hist(predphi.st)
# hist(cos(modvars.accndvi.phist$accndviphi.ts1))
plot(predphi.st, cos(modvars.accndvi.phist$accndviphi.ts1), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.st,cos(modvars.accndvi.phist$accndviphi.ts1))
varimp.phi.st<-varimp(cforest.phi.st)
print(varimp.phi.st[order(varimp.phi.st,decreasing=TRUE)])

partial(cforest.phi.st, pred.var="maxdepth", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="lake_area_ha", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="lake_perim_meters", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="doc", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="cv.accndvi", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="pct.wetlands", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="prcp.normal", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="chla", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="pct.ag", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="tsi.cat", train=modvars.accndvi.phist, type="regression", plot=T)
partial(cforest.phi.st, pred.var="hu4_zoneid", train=modvars.accndvi.phist, type="regression", plot=T)

cforest.phi.lt<-cforest(cos(accndviphi.ts2) ~ lake_area_ha + lake_perim_meters + maxdepth + pct.ag + chla + tsi.cat + hu4_zoneid + 
                          cv.accndvi + pct.wetlands + doc + prcp.normal, 
                        data=modvars.accndvi.philt,controls=cforest_control(ntree=40000))
predphi.lt<-predict(cforest.phi.lt, newdata=modvars.accndvi.philt)
# hist(predphi.lt)
# hist(cos(modvars.accndvi.philt$accndviphi.ts2))
plot(predphi.lt, cos(modvars.accndvi.philt$accndviphi.ts2), xlab="predicted", ylab="empirical", main="cos(phase), short ts",
     xlim=c(-1,1), ylim=c(-1,1))
abline(a=0,b=1)
cor.test(predphi.lt,cos(modvars.accndvi.philt$accndviphi.ts2))
varimp.phi.lt<-varimp(cforest.phi.lt)
print(varimp.phi.lt[order(varimp.phi.lt, decreasing=TRUE)])

partial(cforest.phi.lt, pred.var="hu4_zoneid", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="lake_area_ha", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="prcp.normal", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="pct.wetlands", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="lake_perim_meters", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="doc", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="chla", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="cv.accndvi", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="maxdepth", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="pct.ag", train=modvars.accndvi.philt, type="regression", plot=T)
partial(cforest.phi.lt, pred.var="tsi.cat", train=modvars.accndvi.philt, type="regression", plot=T)
